Fast and scalable Gaussian process modeling with applications to astronomical time series \vspace{-3\baselineskip}

### 2017-03-15 paper/ms.tex
typos and refs

Although we have in mind application of this fast method to stellar variability,
the method is general for one-dimensional GP problems, and may be applied to
other problems.  Accreting black holes show time series which
may be modeled by correlated noise \citep{Kelly:2014};  indeed, this was the
motivation for the original technique developed by Rybicki \& Press \citep{Rybicki:1992,Rybicki:1995}.
This approach may be broadly used for characterizing quasar variability
\citep{MacLeod:2010}, measuring time lags with reverberation mapping
\citep{Zu:2011,Pancoast:2014}, and modeling time delays in multiply-imaged
gravitationally-lensed systems \citep{Press:1998}.

@@

There are three primary applications we envision for the \celerite\ formalism: 1).\ modeling
the variability of an astrophysical system to infer it properties; 2).\
accounting for astrophysical variability as a source of noise when trying to detect
additional phenomena; 3).\ interpolating or extrapolating variability to future or missing
times.  Some examples of the first are determining the characteristic
variability timescale of a quasar \citep{Kelly:2014}, measuring the asteroseismic variation of a star
\citep{Brewer:2009,Corsaro:2014}, measurement of granulation noise \citep{Bastien:2013,Kallinger:2016},
detection of quasi-periodic variability in a high-energy source \citep{McAllister:2016},
or classification of variable objects \citep{Zinn:2016}.  Examples of the
second are detecting transiting exoplanets and correcting for stellar activity in
radial velocity measurements \citep{Haywood:2014,Rajpaul:2015}. Examples of the third are
measuring time delays in
multiply-imaged gravitationally lensed sources \citep{Tewes:2013} and reverberation mapping of AGN.

>>>>

Although we have in mind applications of this fast method to model stellar
variability, the method is general for one-dimensional GP problems, and may be
applied to other problems.
Accreting black holes show time series which may be modeled by correlated
noise \citep{Kelly:2014};  indeed, this was the motivation for the original
technique developed by Rybicki \& Press \citep{Rybicki:1992,Rybicki:1995}.
This approach may be broadly used for characterizing quasar variability
\citep{MacLeod:2010}, measuring time lags with reverberation mapping
\citep{Zu:2011,Pancoast:2014}, and modeling time delays in multiply-imaged
gravitationally-lensed systems \citep{Press:1998}.

@@

There are three primary applications we envision for the \celerite\ formalism:
1).\ modeling the variability of an astrophysical system to infer it
properties; 2).\ accounting for astrophysical variability as a source of noise
when trying to detect additional phenomena; 3).\ interpolating or
extrapolating variability to future or missing times.
Some examples of the first are determining the characteristic variability
timescale of a quasar \citep{Kelly:2014}, measuring the asteroseismic
variation of a star \citep{Brewer:2009,Corsaro:2014}, measurement of
granulation noise \citep{Bastien:2013,Kallinger:2016}, detection of
quasi-periodic variability in a high-energy source \citep{McAllister:2016}, or
classification of variable objects \citep{Zinn:2016}.
Examples of the second are detecting and characterizing transiting exoplanets
\citep[for example,][]{Gibson:2012, Barclay:2015, Evans:2015, Aigrain:2016,
Grunblatt:2016, Luger:2016}
and correcting for stellar activity in radial velocity measurements
\citep[for example,][]{Haywood:2014, Rajpaul:2015}.
Examples of the third are measuring time delays in multiply-imaged
gravitationally lensed sources \citep{Tewes:2013} and reverberation mapping of
AGN.
